---
title: "List and plot new great tit nestboxes"
author: "Nilo Merino Recalde<br>"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  github_document:
    toc: true
    toc_depth: 4
editor_options:
  chunk_output_type: console
---
<br>

This notebook does the following:

* Tells you if there are new great tit nestboxes by checking the google spreadsheets uploaded by fieldworkers
* If there are 10 or more since the last time you checked, then
  + Saves a plot and a list with the new nestboxes and their locations
  + Saves a .gpx file that can be read by mobile gps apps

```{r setup, include=FALSE}
# Knitr settings: 
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

```{r source, include=FALSE}
# Source file
source("./src/viz/plot-new-gretis.R")
```

### Are there new great tit nestboxes?
Always run this first. Do not check more than once every 2 minutes

```{r new, error=TRUE}

# Download data - I started writing this in python 
# and switched to R mid way, so run the first bit by:
py_run_file("./src/read/get-new-gretis-rounds.py")

# Are there at least 10 new nestboxes?
new_boxes_list <- list.files(path = sheet_path,
                             full.names = TRUE,
                             pattern = ".csv")

check_new_nestboxes(new_boxes_list)

```

Run the code below <span style="color:red">only if there ARE new nestboxes</span>.<br>
`stop()` doesn't work with knitr for some reason, so if you do then you'll get duplicates of the previous run.

```{r read-data}

# Prepare data for plotting
newones <- get_new_nestboxes(new_boxes_list)

## This is my google key, remove before sharing
# ggmap::register_google(key = "AIzaSyDfAbhsGQ32byJTmQnpOFjQ_5yr3ViVYeI", write = TRUE)

# Centre the map here
wytham <- c(lon = -1.321898, lat = 51.771250)

# Make subtitles for plot
subtitle <- make_subtitle(new_boxes_list, newones)


# Make plot
new_boxes_map <-
  get_googlemap(
    center = wytham,
    zoom = 14,
    scale = 4,
    maptype = "satellite",
    color = "color"
  ) %>%
  ggmap(base_layer = ggplot(
    aes(x = lon, y = lat), data = newones)) %>%
  +geom_point(colour = "white", size = 2) %>%
  +geom_text_repel(
    aes(label = Nestbox),
    direction = 'both',
    box.padding = 0.40,
    colour = "white",
    segment.alpha = 0.5,
    seed = 30
  ) %>%
  +theme_nothing() %>%
  +ggtitle("New great tit nestboxes", subtitle = subtitle) %>%
  +theme(plot.title = element_text(hjust = 0, size = 30, face = "bold")) %>%
  +theme(plot.subtitle = element_text(hjust = 0, size = 15)) %>%
  +theme(plot.margin = unit(c(1, 1, 1, 1), "cm"))

# Save plot
now <- gsub('.{3}$', '', now(tzone = "UTC"))  %>% sub(" ", "_", .)
filename <- paste0("newboxes_map_", now, ".png")

ggsave(
  new_boxes_map,
  path = figures_path,
  filename = filename,
  width = 30,
  height = 30,
  units = "cm",
  dpi = 350
)

# Make and save list ------------------------------------

list_path <-
  file.path("resources",
            "fieldwork",
            year(today()),
            paste0("new_", now, ".pdf"))

newones %>% select(-x, -y) %>% 
  knitr::kable() %>%  
  kable_styling() %>%
  save_kable(file = list_path)


# Make and save .gpx file --------------------------------

longlats <- data.frame(newones$lon, 
                       newones$lat, 
                       newones$Nestbox)

longlats_spatial <-
  SpatialPointsDataFrame(
    coords = longlats[, c(1, 2)],
    data = longlats,
    proj4string = CRS("+proj=longlat + ellps=WGS84")
  )

writeOGR(
  longlats_spatial,
  dsn = file.path(gpx_path, paste0(now, ".gpx")),
  dataset_options = "GPX_USE_EXTENSIONS=yes",
  layer = "waypoints",
  driver = "GPX",
  overwrite_layer = T
)


 rm(list = ls())

```
